\newpage
\phantomsection % to get the hyperlinks (and bookmarks in PDF) right for index, list of files, bibliography, etc.
\addcontentsline{toc}{chapter}{Abstract}
\begin{abstract}

% In this thesis I tested the performance of a similarity based word alignment method using word embeddings in a zero shot setting on the case of German--Romansh.
% Similarity based word alignments methods using multilingual word embeddings have been shown to perform as well as, or even outperform, statistical word alignment models \autocite{jalili-sabet-etal-2020-simalign}. 
% A clear advantage of using word embeddings for word alignment is that virtually no parallel data is required for first traning an alignment model, since the embeddings are learned from monolingual data.
% In this thesis I tested whether word alignment for sentence pairs in German-Romansh, using multilingual word embeddings, performs as well as statistical models, despite Romansh being an unseen language (it is not part of the data multilingual language models such as mBERT or XLM-R were trained on).
% Using the word embeddings from mBERT for aligning German--Romansh produces word alignemnts that are on par with the word alignments computed by a statistical model (fast\_align). 




Using multilingual word embeddings for computing word alignments has been shown to be competetive with statistical word alignment methods. 
However, the languages on which the experiments were made on were all \enquote{seen} languages, i.e., they were part of the training data for the embeddings. 
In this thesis I show that multilingual word embeddings taken from mBERT can be used for computing word alignments for the \enquote{unseen} language Romansh, aligned against German.
The performance is on par with a baseline statistical model (fast\_align).
I also describe the creation of a gold standard for evaluating the quality of word alignments for German--Romansh, as well as the process of data collection for compiling a trilingual corpus containing press releases in German, Italian and Romansh, published by the Swiss Canton of Grisons. From this corpus, I extracted around 80,000 unique sentence pairs for each language combination. 


\end{abstract}


\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgements}
\section*{Acknowledgements}
First and foremost I would like to profoundly thank Prof.~Dr.~Martin Volk for supervising my thesis and allowing me the right measure of freedom and independence, and also for introducing me to the world of computational linguistics and for being such an inspirational force throughout my studies.

A special thanks also goes to Prof.~Dr.~Rico Sennrich for answering my questions about language models and word alignment during class breaks, as well as to the people at the Department of Computational Linguistics (Dr.~Samuel Läubli, Emma van den Bold,  Phillip Ströbel and Chantal Amrehin to name a few) for their professional, methodological and literature advice. 
Thanks also to Lisa Gasner for allowing me access to her GitHub repository of her German-Romansh corpus.

A big thanks also goes to Steinþór Steingrímsson from the University of Rejkyavik for the long and inspiring conversation at LREC 2022 about word alignment and language technology.

Last but not least I would like to thank my friends and family for supporting me along the way. My most heartfelt gratitude goes to my partner, Mathias Uldack, for putting up with me during the time of my studies. 
I know it wasn't easy.

